### 2020-02-13 thesis/content/chapter-introduction.tex
started with wl section

The field of \ac{ml} on graph-structured data has applications in many domains due to the general expressive power of graphs.
Three common types of graph~\ac{ml} problems are
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Link prediction:}
		A graph with an incomplete edge set is given and the missing edges have to be predicted.
		The generation of friendship suggestions in a social network is a typical example for this.
	\item \textbf{Vertex classification \& regression:}
		Here a class or a score has to be predicted for each vertex of a graph.
		In social graphs this corresponds to the prediction of properties of individuals.
		Another example is the prediction of the amount of traffic at the intersections of a street network.
	\item \textbf{Graph classification \& regression:}
		In this final problem type a single global class or continuous value has to be predicted for an input graph.
		The canonical example for this is the prediction of properties of molecule graphs, e.g.\ the toxicity or solubility of a chemical.
\end{enumerate}
In this thesis we will focus on the last problem type, \ac{gcr}.
A \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the vertices.
Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.

>>>>

The field of \ac{ml} on graph-structured data has applications in many domains due to the general expressive power of graphs.
The three most common types of graph~\ac{ml} problems are
\begin{enumerate}[label=\textbf{\arabic*.}]
	\item \textbf{Link prediction:}
		A graph with an incomplete edge set is given and the missing edges have to be predicted.
		The suggestion of potential friends in a social network is a typical example for this.
	\item \textbf{Vertex classification \& regression:}
		Here a class or a score has to be predicted for each vertex of a graph.
		In social graphs this corresponds to the prediction of properties of individuals, e.g.\ personal preferences or gender.
		Another example is the prediction of the amount of traffic at the intersections of a street network.
	\item \textbf{Graph classification \& regression:}
		In this final problem type a single global class or continuous value has to be predicted for an input graph.
		The canonical example for this is the prediction of properties of molecule graphs, e.g.\ the toxicity or solubility of a chemical.
\end{enumerate}
In this thesis we will focus on the last problem type, \acf{gcr}.
An \ac{ml} method for this problem has to accept graphs of varying size and should be permutation invariant wrt.\ the order in which graph vertices are provided.
Those requirements are not met by the commonly used learners that only accept fixed-size feature vectors as their input, e.g.\ \acp{lm}, \acp{svm} or \acp{mlp}.

### 2020-02-13 thesis/content/chapter-related-work.tex
started with wl section

Since t-norms and t-conorms are commutative and associative they can also be applied to non-empty sets of arbitrary size, i.e. $T(\{ y_1, \dots, y_n \}) = T(y_1, T(\{ y_2, \dots, y_n \}))$ with fixpoint $T(\{ y \}) = y$.
Using this extension, a uninorm $U_\lambda$ can be applied to sets which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
The \ac{lta} disaggregation problem is solved by jointly optimizing a \ac{lm}, i.e.\ the constituent scores ${\{ y_i \in [0, 1] \}}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.

@@

\subsection{Weisfeiler-Lehman Graph Colorings}%
\label{sec:related:character:wl}

The Weisfeiler-Lehman algorithm characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi: \mathcal{V}^k \to \mathcal{C}$ is called a coloring of $G = (\mathcal{V}, \mathcal{E})$.

\paragraph{The 1-dimensional \acs{wl} algorithm}
For the most simple case of $k = 1$, a color is assigned to each vertex of a graph.

>>>>

Since t-norms and t-conorms are commutative and associative they can also be applied to non-empty sets of arbitrary size, i.e. $T(\{ y_1, \dots, y_n \}) = T(y_1, T(\{ y_2, \dots, y_n \}))$ with fixpoint $T(\{ y \}) = y$.
Using this extension, a uninorm $U_\lambda$ can be applied to sets which turns it into a parameterized aggregation function $\mathcal{A}_\lambda: {[0, 1]}^* \to [0, 1]$.
In this simple model the \ac{lta} aggregation problem boils down to the optimization of $\lambda$.
The \ac{lta} disaggregation problem is solved by jointly optimizing a \acf{lm}, i.e.\ the constituent scores ${\{ y_i \in [0, 1] \}}_{c_i \in \bm{c}}$ are described by $y_i = {\left( 1 + \exp(-\theta^\top x_i) \right)}^{-1}$.
Overall an \ac{lta} model is therefore described by the uninorm parameter $\lambda$ and the regression coefficients $\theta$.

@@

\subsection{\acl{wl} Graph Colorings}%
\label{sec:related:character:wl}

The \acf{wl} algorithm~\cite{Weisfeiler1968} characterizes a graph $G = (\mathcal{V}, \mathcal{E})$ by assigning discrete labels $c \in \mathcal{C}$, called \textit{colors}, to vertex $k$-tuples $(v_1, \dots, v_k) \in \mathcal{V}^k$, where $k \in \mathbb{N}$ is the freely choosable \textit{\ac{wl}-dimension}.
A mapping $\chi_{G, k}: \mathcal{V}^k \to \mathcal{C}$ is called a \textit{$k$-coloring} of $G = (\mathcal{V}, \mathcal{E})$.
We say  $\chi'$ \textit{refines} $\chi$ ($\chi' \succeq \chi$) iff.\ $\forall a, b \in \mathcal{V}^k: \chi(a) \neq \chi(b) \rightarrow \chi'(a) \neq \chi'(b)$, i.e.\ $\chi'$ distinguishes at least those tuples that are distinguished by $\chi$.

The $k$-dimensional \ac{wl} algorithm works by iteratively refining $k$-colorings $\chi_{G, k}^0 \preceq \chi_{G, k}^1 \preceq \dots$ of a given graph $G$ until the convergence criterion $\chi_{G, k}^i \equiv \chi_{G, k}^{i+1}$ is satisfied;
the criterion denotes that $\chi_{G, k}^i$ and $\chi_{G, k}^{i+1}$ must be identical up to color/label substitutions.
The final \ac{wl} coloring $\chi_{G, k}$ that is obtained after convergence can be used to quickly check whether two graphs are not isomorphic.
If for two graphs $G$ and $H$ there exists a color $c \in \mathcal{C}$ s.t.\ $\left|\left\{ a \in \mathcal{V}_G^k\, |\, \chi_{G, k}(a) = c \right\}\right|\neq \left|\left\{ b \in \mathcal{V}_H^k\, |\, \chi_{H, k}(b) = c \right\}\right|$, the graphs are $k$-\acs{wl} \textit{distinguishable} ($G \mathrel{{\not\simeq}_k} H$) and therefore non-isomorphic ($G \not\simeq H$). % chktex 21
The opposite does however not necessarily hold;
two $k$-\acs{wl} indistinguishable graphs are not always isomorphic, i.e.\ $G \mathrel{{\simeq}_k} H \centernot\implies G \simeq H$.
To understand which types of graphs are $k$-\acs{wl} distinguishable, we will proceed in three steps:
\begin{enumerate}
	\item First the color refinement algorithm for the most simple case of $k = 1$ is described.
	\item Then we extend the definitions and intuitions from the 1-dimensional \ac{wl} algorithm to its higher-dimensional generalization.
	\item Lastly we will discuss the discriminative power of the \acs{wl} algorithm and its relation to the \acs{wl}-dimension $k$.
\end{enumerate}

\subsubsection{The 1-dimensional \acs{wl} algorithm}
In the 1-dimensional \ac{wl} algorithm, a color is assigned to each vertex of a graph.
If the vertices $\mathcal{V}_G$ of the input graph $G$ are already labeled via some labeling $l_G: \mathcal{V}_G \to L$, those labels can be used as the initial graph coloring $\chi_{G,1}^0 = l_G$.
For unlabeled graphs a constant coloring is used instead, i.e.\ $\forall v \in \mathcal{V}_G: \chi_{G,1}^0(v) = \texttt{A}$ for some initial color $\texttt{A} \in \mathcal{C}$.
In each iteration of the 1-\acs{wl} color refinement algorithm, the following neighborhood aggregation scheme is used to compute a new color for all vertices:
\begin{align}
	\chi_{G,1}^{i+1}(v) := h\left(\chi_{G,1}^{i}(v), \{ \chi_{G,1}^{i}(u) | u \in \Gamma_{G,1}(v) \}\right)
\end{align}
